Comparison of Catalogs

### 2024-07-19 compare/compare.qmd
Fixed errordars in scatter

```{python}
import numpy as np
import matplotlib.pyplot as plt
from sklearn.metrics import r2_score
from astropy.io import ascii
from astropy.coordinates import SkyCoord
from astropy.table import Table, join, QTable
import astropy.units as u
from astropy.visualization import quantity_support
#quantity_support()
from astropy.modeling import models, fitting
from scipy.stats import pearsonr
import glob
import os
plt.style.use('ggplot')

@@

1.  [$R^2$]{.underline}: Measures the proportion of variance explained by the linear model.
2.  [Slope of the Fitted Line]{.underline}: Should be close to 1 for a 1-1 correlation.

@@

    # Calculate R-squared
    r2 = r2_score(y_data, y_pred)

    # Calculate Pearson correlation coefficient
    corr, _ = pearsonr(x_data, y_data)

    return slope, intercept, r2, corr

def scatter_plot(x, y, xerr =0, yerr = 0):
    """
    Generates a scatter plot of two datasets with optional error bars, fits a linear model to the data, 
    and displays the fitted line on the plot.

@@

@@

    # Plot the data with error bars
    # errorbar -> needs np.ma -> no units
    plt.errorbar(x_data, y_data, xerr, yerr, alpha = 0.3, c = "blue", label = "Data", fmt=".k")

    # Plot the fitted line with unit support
    plt.plot(x, c*x + m*x.unit, c = "red", label = "Fit")

@@

|  LCV  | HECATE |   Description   | Linear Correlation |
|:-----:|:------:|:---------------:|:------------------:|
| Ra_1  |  RA_2  | Right Ascension |                    |
| Dec_1 | DEC_2  |   Declination   |                    |
|  Dis  |   D    |    Distance     |                    |

@@

|        LCV        |        HECATE         |               Description               | Linear Correlation |
|:----------------:|:----------------:|:-----------------:|:----------------:|
| W50 (with errors) |                       | H I line with at 50% level from maximum |                    |
|       RVel        |    V (with errors)    |      Heliocentric radial velocity       |                    |
|        Vm         |                       |    Amplitude of rotational velocity     |                    |
|        VLG        | V/V_VIR (with errors) |             Radial velocity             |                    |
| cz (with errors)  |                       |          Heliocentric velocity          |                    |

@@

|      LCV      |        HECATE        |                        Description                        | Linear Correlation |
|:----------------:|:----------------:|:-----------------:|:----------------:|
|     TType     |   T (with errors)    | Numerical Hubble type following the de Vaucouleurs system |                    |
|      inc      |          I           |                     Inclination (deg)                     |                    |
| a26_1 (Major) | R1 (Semi-major axis) |                 angular diameter (arcmin)                 |                    |

>>>>

```{python}
#| echo: false
import numpy as np
import matplotlib.pyplot as plt
from sklearn.metrics import r2_score
from astropy.io import ascii
from astropy.coordinates import SkyCoord
from astropy.table import Table, join, QTable
import astropy.units as u
from astropy.visualization import quantity_support
#quantity_support()
from astropy.modeling import models, fitting
from scipy.stats import pearsonr
import glob
import os
plt.style.use('ggplot')

@@

1.  [$R^2$]{.underline}: Measures the proportion of variance explained by the linear model.
2.  [Slope of the Fitted Line]{.underline}: Should be close to 1 for a 1-1 correlation.
3.  [Pearson Correlation $\rho$]{.underline}: Measures the strength and direction of the linear relationship between two variables, ranging from -1 to 1.

In simple linear regression, $R^2$ is the square of the Pearson correlation coefficient $r$.

@@

    # Remove NaN values
    mask = ~np.isnan(y_data) & ~np.isnan(y_pred)
    y_data_clean = y_data[mask]
    y_pred_clean = y_pred[mask]
    mask = ~np.isnan(x_data) 
    x_data_clean = x_data[mask]

    # Calculate R-squared
    r2 = r2_score(y_data_clean, y_pred_clean)

    # Calculate Pearson correlation coefficient
    corr = np.sqrt(np.abs(r2))

    return slope, intercept, r2, corr
```

4.  [Scatter plots]{.underline}: Scatter plots are essential for visually assessing the relationship between two datasets, identifying correlations, trends, and outliers, and evaluating the fit of linear models.

```{python}
def scatter_plot(x, y, xerr = None, yerr = None):
    """
    Generates a scatter plot of two datasets with optional error bars, fits a linear model to the data, 
    and displays the fitted line on the plot.

@@

    # Handle default values for xerr and yerr
    if xerr is None:
        xerr_d = np.ma.zeros(len(x_data))
    else:
        xerr_d = np.ma.array(xerr.value, mask=np.isnan(xerr))
    if yerr is None:
        yerr_d = np.ma.zeros(len(y_data))
    else:
        yerr_d = np.ma.array(yerr.value, mask=np.isnan(yerr))

@@

    # Plot the data with error bars
    # errorbar -> needs np.ma -> no units
    plt.errorbar(x_data, y_data, xerr = xerr_d, yerr = yerr_d, alpha = 0.3, c = "blue", label = "Data", fmt=".")

    # Plot the fitted line with unit support
    plt.plot(x, c*x + m*x.unit, c = "red", label = f"Fit: {y.info.name}$_H $= {c:.2f}$\cdot${x.info.name}$_L${m:+.2f}\n $R^2=$ {r2:.2f}")

@@

```{python}
#| echo: false
ra_corr = round(compare_data(dt["Ra_1"],dt["RA_2"])[3],3)
dec_corr = round(compare_data(dt["Dec_1"],dt["DEC_2"])[3],3)
d_corr = round(compare_data(dt["Dis"],dt["D"])[3],3)
```

|  LCV  | HECATE |   Description   | Pearson Correlation \[-1,1\] |
|:-----:|:------:|:---------------:|:----------------------------:|
| Ra_1  |  RA_2  | Right Ascension |      `{python} ra_corr`      |
| Dec_1 | DEC_2  |   Declination   |     `{python} dec_corr`      |
|  Dis  |   D    |    Distance     |      `{python} d_corr`       |

```{python}

scatter_plot(dt["Ra_1"], dt["RA_2"]), scatter_plot(dt["Dec_1"], dt["DEC_2"]), scatter_plot(dt["Dis"], dt["D"])
```

@@

```{python}
rvel_corr = round(compare_data(dt["RVel"],dt["V"])[3],3)
rvel_corr_v = round(compare_data(dt["RVel"],dt["V_VIR"])[3],3)
vlg_corr = round(compare_data(dt["VLG"],dt["V"])[3],3)
vlg_corr_v = round(compare_data(dt["VLG"],dt["V_VIR"])[3],3)
cz_corr = round(compare_data(dt["cz"],dt["V"])[3],3)
cz_corr_v = round(compare_data(dt["cz"],dt["V_VIR"])[3],3)
```

|       LCV        |        HECATE         |         Description          |        Linear Correlation         |
|:----------------:|:------------------:|:----------------:|:----------------:|
|       RVel       | V/V_VIR (with errors) | Heliocentric radial velocity | `{python} rvel_corr, rvel_corr_v` |
|       VLG        | V/V_VIR (with errors) |       Radial velocity        |  `{python} vlg_corr, vlg_corr_v`  |
| cz (with errors) | V/V_VIR (with errors) |    Heliocentric velocity     |   `{python} cz_corr, cz_corr_v`   |

```{python}
scatter_plot(dt["RVel"], dt["V"], yerr = dt["E_V"])
```

@@

```{python}
#| echo: false

ttype_corr = round(compare_data(dt["TType"], dt["T"])[3], 4)
inc_corr = round(compare_data(dt["inc"], dt["I"])[3], 3)
#a26_corr = round(compare_data(dt["a26_1"], dt["R1"])[3], 3)
#it doesnt even compute
```

|      LCV      |        HECATE        |                        Description                        | Pearson Correlation \[-1,1\] |
|:---------------:|:---------------:|:---------------------:|:---------------:|
|     TType     |   T (with errors)    | Numerical Hubble type following the de Vaucouleurs system |    `{python} ttype_corr`     |
|      inc      |          I           |                     Inclination (deg)                     |     `{python} inc_corr`      |
| a26_1 (Major) | R1 (Semi-major axis) |                 angular diameter (arcmin)                 |              0               |

As we can see we don't have any significant correlation between the data

```{python}

dt["TType"].info, dt["T"].info
```
